### 2022-02-28 generated-example-data.tex

\def\countFields{32}
\noindent\texttt{
\hspace*{\codeIndent}\{\hspace*{\codeIndent}\newline
\hspace*{2\codeIndent} accessed: "14 July 2020",\newline
\hspace*{2\codeIndent} doubleChecked: "17 January 2021",\newline
\hspace*{2\codeIndent} authors: "Callahan A, Steinberg E, Fries JA, Gomba \ldots,\newline
\hspace*{2\codeIndent} year: 2020,\newline
\hspace*{2\codeIndent} title: "Estimating the efficacy of symptom-based \ldots,\newline
\hspace*{2\codeIndent} volume: 3,\newline
\hspace*{2\codeIndent} number: 95,\newline
\hspace*{2\codeIndent} journal: "Nature Digital Medicine",\newline
\hspace*{2\codeIndent} doi: "10.1038/s41746-020-0300-0",\newline
\hspace*{2\codeIndent} dataComment: "On request",\newline
\hspace*{2\codeIndent} hasCodeInPrinciple: 1,\newline
\hspace*{2\codeIndent} codeComment: "``Code is available upon request from th \ldots,\newline
\hspace*{2\codeIndent} pages: 3\newline
\hspace*{\codeIndent}\}
}