### 2022-02-28 generated-info-for-main.tex

\newlabel{supplementary-material-makefiles}{{2}{40}{DUMMYA}{DUMMYB}{DUMMYC}}
\newlabel{on-code-data-publication}{{3}{43}{XXX}{YYYY}{ZZZZ}}
\newlabel{supplementary-Speigelhalter-section}{{4}{46}{DUMMYA}{DUMMYB}{DUMMYC}}
\def\MaxMainPaperCitationNumber{79}
\newlabel{supplementary-journal-policies-section}{{5}{55}{DUMMYA}{DUMMYB}{DUMMYC}}
\newlabel{supplementary-summary-table}{{5}{60}{DUMMY1}{DUMMY2}{DUMMY3}}
\expandafter\def \csname cite-MetricSelectionFramework\endcsname{[113]}
\expandafter\def \csname cite-PENet\endcsname{[117]}
\expandafter\def \csname cite-PostoperativeOutcomes.RiskNet\endcsname{[119]}
\expandafter\def \csname cite-philter-ucsf\endcsname{[120]}
\expandafter\def \csname cite-AI-CDSS-Cardiovascular-Silo\endcsname{[121]}
\expandafter\def \csname cite-SiameseChange\endcsname{[123]}
\expandafter\def \csname cite-dryad.1g1jwstrw\endcsname{[124]}
\expandafter\def \csname cite-rsos.192210\endcsname{[125]}
\expandafter\def \csname cite-ichHKrWj7hqlznOaR6NQVzITgp40dlqWvWAgAxyafiQ\endcsname{[126]}
\expandafter\def \csname cite-rsos.200566\endcsname{[127]}
\expandafter\def \csname cite-?view.only=87ae173f775b40d79d6cd0fdcf6d4a9c\endcsname{[128]}
\expandafter\def \csname cite-dryad.vx0k6djnr\endcsname{[131]}
\expandafter\def \csname cite-lactModel\endcsname{[132]}
\expandafter\def \csname cite-LRM\endcsname{[134]}
\expandafter\def \csname cite-1\endcsname{[135]}
\expandafter\def \csname cite-manifold-ga\endcsname{[140]}
\expandafter\def \csname cite-blast-ct\endcsname{[142]}