### 2022-02-28 generated-summary-table.tex

\begin{tabular}{|rrrc|}\hline
Number of papers sampled relying on code&32&100\%&\\\hline\hline
\multicolumn{4}{|l|}{\textbf{Access to code}}\\
Have some or all code available&12&38\%&\\
Some or all code in principle available on request&8&25\%&\\
No code available&12&38\%&\\\hline
\multicolumn{4}{|l|}{{\textbf{Evidence of basic good software engineering practice}}}\\
Evidence program designed rigorously&{0}&{0\%}&\\
Evidence source code properly tested&{0}&{0\%}&\\
\begin{change}Evidence of any tool-based development\end{change}&{\begin{change}0\end{change}}&{\begin{change}0\%\end{change}}&\\
\begin{change}Team or open source based development\end{change}&{\begin{change}0\end{change}}&{\begin{change}0\%\end{change}}&\\
Other methods, e.g., independent coding methods&{1}&{3\%}&\\\hline
\multicolumn{4}{|l|}{\textbf{Documentation and comments}}\\
Substantial code documentation and comments&2&6\%&\\
Comments explain some code intent&3&9\%&\\
Procedural comments (e.g., author, date, copyright)&10&31\%&\\
No usable comments&17&53\%&\\\hline
\multicolumn{4}{|l|}{\textbf{Repository use}}\\
Code repository (e.g., GitHub) --- 1 was empty&10&31\%&\\
Data repository (e.g., Dryad or GitHub)&9&28\%&\\\hline
\multicolumn{4}{|l|}{\textbf{Adherence to journal code policy (if any)}}\\
Papers published in journals with code policies&26&81\%&\\
Clear breaches of code policy (if any) & 11&42\%&($N=26$)\\
\hline\end{tabular}