### 2022-02-28 generated-supplementary-references.tex

\vbox{\bibitem{ref-1}
Callahan A,  Steinberg E,  Fries JA,  Gombar S,  Patel B,  Corbin CK and  Shah NH, ``Estimating the efficacy of symptom-based screening for COVID-19,'' \emph{Nature Digital Medicine}, \textbf{3}(95):3pp, 2020. DOI \texttt{10.1038/s41746-020-0300-0}\\\hfill{Accessed 14 July 2020.}\ {Double-checked 17 January 2021}.}\bibskip

\vbox{\maprepo{MetricSelectionFramework}\bibitem{ref-2}
Kanzler CM,  Rinderknecht MD,  Schwarz A,  Lamers I,  Gagnon C,  Held JPO,  Feys P,  Luft AR,  Gassert R and  Lambercy O, ``A data-driven framework for selecting and validating digital health metrics: use-case in neurological sensorimotor impairments,'' \emph{Nature Digital Medicine}, \textbf{3}(80):17pp, 2020. DOI \texttt{10.1038/s41746-020-0286-7} {Code \url{github.com/ChristophKanzler/MetricSelectionFramework}}\\\hfill{Accessed 14 July 2020.}\ {Double-checked 17 January 2021}.}\bibskip

\vbox{\bibitem{ref-3}
Roy A,  Nikolitch K,  McGinn R,  Jinah S,  Klement W and  Kaminsky ZA, ``A machine learning approach predicts future risk to suicidal ideation from social media data,'' \emph{Nature Digital Medicine}, \textbf{3}(78):12pp, 2020. DOI \texttt{10.1038/s41746-020-0287-6}\\\hfill{Accessed 14 July 2020.}\ {Double-checked 17 January 2021}.}\bibskip

\vbox{\bibitem{ref-4}
Levine DM,  Co Z,  Newmark LP,  Groisser AR,  Holmgren AJ,  Haas JA and  Bates DW, ``Design and testing of a mobile health application rating tool,'' \emph{Nature Digital Medicine}, \textbf{3}(74):7pp, 2020. DOI \texttt{10.1038/s41746-020-0268-9}\\\hfill{Accessed 14 July 2020.}\ {Double-checked 17 January 2021}.}\bibskip

\vbox{\bibitem{ref-5}
Kannampallil T,  Smyth JM,  Jones S,  Payne PRO and  Ma J, ``Cognitive plausibility in voice-based AI health counselors,'' \emph{Nature Digital Medicine}, \textbf{3}(72):4pp, 2020. DOI \texttt{10.1038/s41746-020-0278-7}\\\hfill{Accessed 14 July 2020.}\ {Double-checked 17 January 2021}.}\bibskip

\vbox{\maprepo{PENet}\bibitem{ref-6}
Huang S,  Kothari T,  Banerjee I,   Chute C,  Ball RL,  Borus N,  Huang A,  Patel BN,  Rajpurkar P,  Irvin J,  Dunnmon J,   Bledsoe J,  Shpanskaya K,  Dhaliwal A,  Zamanian R,  Ng AY and  Lungren MP, ``PENet a scalable deep-learning model for automated diagnosis of pulmonary embolism using volumetric CT imaging,'' \emph{Nature Digital Medicine}, \textbf{3}(61):9pp, 2020. DOI \texttt{10.1038/s41746-020-0266-y} {Code \url{github.com/marshuang80/PENet}}\\\hfill{Accessed 14 July 2020.}\ {Double-checked 17 January 2021}.}\bibskip

\vbox{\bibitem{ref-7}
Dhruva SS,  Ross JS,  Akar JG,   Caldwell B,  Childers K,  Chow W,  Ciaccio L,  Coplan P,  Dong J,  Dykhoff HJ,  Johnston S,  Kellogg T,  Long C,  Noseworthy PA,  Roberts K,  Saha A,  Yoo A and  Shah ND, ``Aggregating multiple real-world data sources using a patient-centered health-data-sharing platform,'' \emph{Nature Digital Medicine}, \textbf{3}(60):9pp, 2020. DOI \texttt{10.1038/s41746-020-0265-z}\\\hfill{Accessed 14 July 2020.}\ {Double-checked 17 January 2021}.}\bibskip

\vbox{\maprepo{PostoperativeOutcomes.RiskNet}\bibitem{ref-8}
Hofer IS,  Lee C,  Gabel E,  Baldi P and  Cannesson M, ``Development and validation of a deep neural network model to predict postoperative mortality, acute kidney injury, and reintubation using a single feature set,'' \emph{Nature Digital Medicine}, \textbf{3}(58):10pp, 2020. DOI \texttt{10.1038/s41746-020-0248-0} {Code \url{github.com/cklee219/PostoperativeOutcomes_RiskNet}}\\\hfill{Accessed 14 July 2020.}\ {Double-checked 17 January 2021}.}\bibskip

\vbox{\maprepo{philter-ucsf}\bibitem{ref-9}
Norgeot B,  Muenzen K,  Peterson TA,  Fan X,  Glicksberg BS,  Schenk G,  Rutenberg E,  Oskotsky B,  Sirota M,  Yazdany J,  Schmajuk G,  Ludwig D,  Goldstein T and  Butte AJ, ``Protected Health Information filter (Philter): accurately and securely de-identifying free-text clinical notes,'' \emph{Nature Digital Medicine}, \textbf{3}(57):8pp, 2020. DOI \texttt{10.1038/s41746-020-0258-y} {Code \url{github.com/BCHSI/philter-ucsf}}\\\hfill{Accessed 14 July 2020.}\ {Double-checked 17 January 2021}.}\bibskip

\vbox{\maprepo{AI-CDSS-Cardiovascular-Silo}\bibitem{ref-10}
Choi D,  Park JJ,  Ali T and  Lee S, ``Artificial intelligence for the diagnosis of heart failure,'' \emph{Nature Digital Medicine}, \textbf{3}(54):6pp, 2020. DOI \texttt{10.1038/s41746-020-0261-3} {Code \url{github.com/ubiquitous-computing-lab/AI-CDSS-Cardiovascular-Silo}}\\\hfill{Accessed 14 July 2020.}\ {Double-checked 17 January 2021}.}\bibskip

\vbox{\bibitem{ref-11}
Hilton CB,  Milinovich A,  Felix C,  Vakharia N,   Crone T,   Donovan C,  Proctor A and  Nazha A, ``Personalized predictions of patient outcomes during and after hospitalization using artificial intelligence,'' \emph{Nature Digital Medicine}, \textbf{3}(51):8pp, 2020. DOI \texttt{10.1038/s41746-020-0249-z}\\\hfill{Accessed 14 July 2020.}\ {Double-checked 17 January 2021}.}\bibskip

\vbox{\maprepo{SiameseChange}\bibitem{ref-12}
Li MD,  Chang K,  Bearce B,  Chang BY,  Huang AJ,  Campbell JP,  Brown JM,  Singh P,  Hoebel KV,   Erdo{\u g}mu\c{s} D,  Ioannidis S,  Palmer W,  Chiang MF and  Kalpathy-Cramer J, ``Siamese neural networks for continuous disease severity evaluation and change detection in medical imaging,'' \emph{Nature Digital Medicine}, \textbf{3}(48):9pp, 2020. DOI \texttt{10.1038/s41746-020-0255-1} {Code \url{github.com/QTIM-Lab/SiameseChange}}\\\hfill{Accessed 14 July 2020.}\ {Double-checked 19 January 2021}.}\bibskip

\vbox{\maprepo{dryad.1g1jwstrw}\bibitem{paper-usesRMarkdown}
Hoffman JI,  Nagel R,  Litzke V,  Wells DA and  Amos W, ``Genetic analysis of \emph{Boletus edulis\/} suggests that intra-specific competition may reduce local genetic diversity as a woodland ages,'' \emph{Royal Society Open Science}, \textbf{7}(200419):13pp, 2020. DOI \texttt{10.1098/rsos.200419} {Code \url{datadryad.org/stash/dataset/doi:10.5061/dryad.1g1jwstrw}}\\\hfill{Accessed 22 July 2020.}\ {Double-checked 26 January 2021}.}\bibskip

\vbox{\maprepo{rsos.192210}\bibitem{ref-14}
Gr\"onquist P,  Panchadcharam P,  Wood D,  Menges A,  R\"uggeberg M and  Wittel FK, ``Computational analysis of hygromorphic self-shaping wood gridshell structures,'' \emph{Royal Society Open Science}, \textbf{7}(192210):9pp, 2020. DOI \texttt{10.1098/rsos.192210} {Code \url{royalsocietypublishing.org/doi/suppl/10.1098/rsos.192210}}\\\hfill{Accessed 22 July 2020.}\ {Double-checked 26 January 2021}.}\bibskip

\vbox{\maprepo{ichHKrWj7hqlznOaR6NQVzITgp40dlqWvWAgAxyafiQ}\bibitem{onlyPaperWithChecks}
Amos W, ``Signals interpreted as archaic introgression appear to be driven primarily by faster evolution in Africa,'' \emph{Royal Society Open Science}, \textbf{7}(191900):9pp, 2020. DOI \texttt{10.1098/rsos.191900} {Code \url{datadryad.org/stash/share/ichHKrWj7hqlznOaR6NQVzITgp40dlqWvWAgAxyafiQ}}\\\hfill{Accessed 22 July 2020.}\ {Double-checked 26 January 2021}.}\bibskip

\vbox{\maprepo{rsos.200566}\bibitem{ref-16}
Gordon M,  Viganola D,  Bishop M,  Chen Y,  Dreber A,  Goldfedder B,  Holzmeister F,  Johannesson M,  Liu Y,  Twardy C,  Wang J and  Pfeiffer T, ``Are replication rates the same across academic fields? Community forecasts from the DARPA SCORE programme,'' \emph{Royal Society Open Science}, \textbf{7}(200566):7pp, 2020. DOI \texttt{10.1098/rsos.200566} {Code \url{royalsocietypublishing.org/doi/suppl/10.1098/rsos.200566}}\\\hfill{Accessed 22 July 2020.}\ {Double-checked 26 January 2021}.}\bibskip

\vbox{\maprepo{?view.only=87ae173f775b40d79d6cd0fdcf6d4a9c}\bibitem{ref-17}
Evans D and  Field AP, ``Predictors of mathematical attainment trajectories across the primary-to-secondary education transition: parental factors and the home environment,'' \emph{Royal Society Open Science}, \textbf{7}(200422):20pp, 2020. DOI \texttt{10.1098/rsos.200422} {Code \url{osf.io/a5xsz/?view_only=87ae173f775b40d79d6cd0fdcf6d4a9c}}\\\hfill{Accessed 22 July 2020.}\ {Double-checked 26 January 2021}.}\bibskip

\vbox{\bibitem{ethics-paper}
Beale N,  Battey H,  Davison AC and  MacKay RS, ``An unethical optimization principle,'' \emph{Royal Society Open Science}, \textbf{7}(200462):11pp, 2020. DOI \texttt{10.1098/rsos.200462}\\\hfill{Accessed 22 July 2020.}\ {Double-checked 26 January 2021}.}\bibskip

\vbox{\bibitem{ref-19}
Cherevko AA,  Gologush TS,  Petrenko IA,  Ostapenko VV and  Panarin VA, ``Modelling of the arteriovenous malformation embolization optimal scenario,'' \emph{Royal Society Open Science}, \textbf{7}(191992):16pp, 2020. DOI \texttt{10.1098/rsos.191992}\\\hfill{Accessed 22 July 2020.}\ {Double-checked 26 January 2021}.}\bibskip

\vbox{\maprepo{dryad.vx0k6djnr}\bibitem{ref-20}
Soczawa-Stronczyk AA and  Bocian M, ``Gait coordination in overground walking with a virtual reality avatar,'' \emph{Royal Society Open Science}, \textbf{7}(200622):19pp, 2020. DOI \texttt{10.1098/rsos.200622} {Code \url{datadryad.org/stash/dataset/doi:10.5061/dryad.vx0k6djnr}}\\\hfill{Accessed 22 July 2020.}\ {Double-checked 26 January 2021}.}\bibskip

\vbox{\maprepo{lactModel}\bibitem{ref-21}
Duruz S,  Vajana E,  Burren A,  Flury C and  Joost S, ``Big dairy data to unravel effects of environmental, physiological and morphological factors on milk production of mountain-pastured Braunvieh cows,'' \emph{Royal Society Open Science}, \textbf{7}(200638):13pp, 2020. DOI \texttt{10.1098/rsos.200638} {Code \url{github.com/SolangeD/lactModel}}\\\hfill{Accessed 22 July 2020.}\ {Double-checked 26 January 2021}.}\bibskip

\vbox{\bibitem{ref-22}
de Azevedo EZD,  Dantas DV and  Daura-Jorge FG, ``Risk tolerance and control perception in a game-theoretic bioeconomic model for small-scale fisheries,'' \emph{Royal Society Open Science}, \textbf{7}(200621):11pp, 2020. DOI \texttt{10.1098/rsos.200621}\\\hfill{Accessed 22 July 2020.}\ {Double-checked 26 January 2021}.}\bibskip

\vbox{\maprepo{LRM}\bibitem{ref-23}
Abdolhosseini-Qomi AM,  Jafari SH,  Taghizadeh A,  Yazdani N,  Asadpour M and  Rahgozar M, ``Link prediction in real-world multiplex networks via layer reconstruction method,'' \emph{Royal Society Open Science}, \textbf{7}(191928):22pp, 2020. DOI \texttt{10.1098/rsos.191928} {Code \url{github.com/UT-NSG/LRM}}\\\hfill{Accessed 22 July 2020.}\ {Double-checked 26 January 2021}.}\bibskip

\vbox{\maprepo{1}\bibitem{example-numerical-error}
Webster J and  Amos M, ``A Turing test for crowds,'' \emph{Royal Society Open Science}, \textbf{7}(200307):12pp, 2020. DOI \texttt{10.1098/rsos.200307} {Code \url{figshare.com/collections/Supplementary_information_for_Webster_J_and_Amos_M_A_Turing_Test_for_Crowds_/4859118/1}}\\\hfill{Accessed 22 July 2020.}\ {Double-checked 26 January 2021}.}\bibskip

\vbox{\bibitem{ref-25}
Zhu Y-l,  Wang C-J,  Gao F,  Xiao Z-x,  Zhao P-l and  Wang J-y, ``Calculation on surface energy and electronic properties of CoS${}_2$,'' \emph{Royal Society Open Science}, \textbf{7}(191653):12pp, 2020. DOI \texttt{10.1098/rsos.191653}\\\hfill{Accessed 22 July 2020.}\ {Double-checked 26 January 2021}.}\bibskip

\vbox{\bibitem{ref-26}
Yu B,  Scott CJ,  Xue X,  Yue X and  Dou X, ``Derivation of global ionospheric Sporadic E critical frequency ($f_o$Es) data from the amplitude variations in GPS/GNSS radio occultations,'' \emph{Royal Society Open Science}, \textbf{7}(200320):15pp, 2020. DOI \texttt{10.1098/rsos.200320}\\\hfill{Accessed 22 July 2020.}\ {Double-checked 26 January 2021}.}\bibskip

\vbox{\bibitem{ref-27}
Joon-myoung K,  Younghoon C,  Ki-Hyun J,  Soohyun C,  Kyung-Hee K,  Seung D B,  Soomin J,  Jinsik P and  Byung-Hee O, ``A deep learning algorithm to detect anaemia with ECGs: a retrospective, multicentre study,'' \emph{Lancet Digital Health}, \textbf{2}(7):9pp, 2020. DOI \texttt{10.1016/S2589-7500(20)30108-4}\\\hfill{Accessed 24 July 2020.}\ {Double-checked 26 January 2021}.}\bibskip

\vbox{\bibitem{ref-28}
Zhu H,  Cheng C,  Yin H,  Li X,  Zuo P,  Ding J,  Lin F,  Wang J,  Zhou B,  Li Y,  Hu S,  Xiong Y,  Wang B,  Wan G,  Yang X and  Yuan Y, ``Automatic multilabel electrocardiogram diagnosis of heart rhythm or conduction abnormalities with deep learning: a cohort study,'' \emph{Lancet Digital Health}, \textbf{2}(7):9pp, 2020. DOI \texttt{10.1016/S2589-7500(20)30107-2}\\\hfill{Accessed 24 July 2020.}\ {Double-checked 26 January 2021}.}\bibskip

\vbox{\maprepo{manifold-ga}\bibitem{ref-29}
Fung R,  Villar J,  Dashti A,  Ismail LC,  Staines-Urias E,  Ohuma EO,  Salomon LJ,  Victora CG,  Barros FC,  Lambert A,  Carvalho M,  Jaffer Y A,  Noble JA,  Gravett MG,  Purwar M,  Pang R,  Bertino E,  Munim S,  Min AM,  McGready R,  Norris SA,  Bhutta ZA,  Kennedy SH,  Papageorghiou AT and  Ourmazd A, ``Achieving accurate estimates of fetal gestational age and personalised predictions of fetal growth based on data from an international prospective cohort study: a population-based machine learning study,'' \emph{Lancet Digital Health}, \textbf{2}(7):7pp, 2020. DOI \texttt{10.1016/S2589-7500(20)30131-X} {Code \url{github.com/ki-analysis/manifold-ga}}\\\hfill{Accessed 24 July 2020.}\ {Double-checked 26 January 2021}.}\bibskip

\vbox{\bibitem{ref-30}
Sabanayagam C,  Xu D,  Ting DSW,  Nusinovici S,  Banu R,  Hamzah H,  Lim C,  Tham Y-C,  Cheung CY,  Tai ES,  Wang XY,  Jonas JB,   Cheng C-Y,  Lee ML,  Hsu W and  Wong TY, ``A deep learning algorithm to detect chronic kidney disease from retinal photographs in community-based populations,'' \emph{Lancet Digital Health}, \textbf{2}(7):7pp, 2020. DOI \texttt{10.1016/S2589-7500(20)30063-7}\\\hfill{Accessed 24 July 2020.}\ {Double-checked 26 January 2021}.}\bibskip

\vbox{\maprepo{blast-ct}\bibitem{ref-31}
Monteiro M,  Newcombe VF,  Mathieu F,  Adatia K,  Kamnitsas K,  Ferrante E,  Das T,  Whitehouse D,  Rueckert D,  Menon DK and  Glocker B, ``Multiclass semantic segmentation and quantification of traumatic brain injury lesions on head CT using deep learning: an algorithm development and multicentre validation study,'' \emph{Lancet Digital Health}, \textbf{2}(7):8pp, 2020. DOI \texttt{10.1016/S2589-7500(20)30085-6} {Code \url{github.com/biomedia-mira/blast-ct}}\\\hfill{Accessed 24 July 2020.}\ {Double-checked 27 January 2021}.}\bibskip

\vbox{\bibitem{ref-32}
Liu K-L,  Wu T,  Chen P-T,  Tsai Y M,  Roth H,  Wu M-S,  Liao W-C and  Wang W, ``Deep learning to distinguish pancreatic cancer tissue from non-cancerous pancreatic tissue: a retrospective study with cross-racial external validation,'' \emph{Lancet Digital Health}, \textbf{2}(7):10pp, 2020. DOI \texttt{10.1016/S2589-7500(20)30078-9}\\\hfill{Accessed 24 July 2020.}\ {Double-checked 27 January 2021}.}\bibskip