### 2023-01-31 generated/info-for-main.tex
Updating to January 2023 submission to Computer Journal

\newlabel{supplementary-material-makefiles}{{11}{38}{DUMMYA}{DUMMYB}{DUMMYC}}
\newlabel{on-code-data-publication}{{12}{41}{DUMMYA}{DUMMYB}{DUMMYC}}
\newlabel{RAP-diagrams}{{12}{44}{DUMMYA}{DUMMYB}{DUMMYC}}
\newlabel{excelFigure}{{2}{44}{AAA}{BBB}{CCC}}
\newlabel{supplementary-Spiegelhalter-section}{{13}{44}{DUMMYA}{DUMMYB}{DUMMYC}}
\newlabel{survey-section}{{14}{47}{DUMMYA}{DUMMYB}{DUMMYC}}
\newlabel{supplementaryTableRepoSummary}{{4}{49}{DUMMYA}{DUMMYB}{DUMMYC}}
\newlabel{supplementary-journal-policies-section}{{14}{52}{DUMMYA}{DUMMYB}{DUMMYC}}
\newlabel{restrictive-policies}{{14}{56}{DUMMYA}{DUMMYB}{DUMMYC}}
\def\MaxMainPaperCitationNumber{121}