### 2023-01-31 generated/legend.tex

\begin{tabular}{lp{4.5in}}
\flagStyle{$\sf P_c$}&Journal has a code policy (see section~\ref{supplementary-journal-policies-section})\\
\flagStyle{$\sf P_{{\mbox{\scriptsize c-breach}}}$}&Paper breaches journal code policy (see section~\ref{supplementary-journal-policies-section})\\
\flagStyle{$\sf R_c$}&Paper uses a code repository (e.g., GitHub)\\
\flagStyle{$\sf R_{\mbox{\scriptsize c-empty}}$}&Code repository contains no code\\
\flagStyle{$\sf R_d$}&Paper uses a data repository (e.g., Dryad, Figshare, GitHub)\\
\flagStyle{$\sf S_{\mbox{\scriptsize NONE}}$}&No code available at all (note: code is not expected for standard models, systems or statistical methods)\\
\flagStyle{$\sf S_p$}&Paper says source code is available in principle\\
\flagStyle{$\sf S_{+}$}&Paper or URL provides source code\\
\flagStyle{$\sf S_{{\mbox{\scriptsize rigorous}}}$}&Evidence that source code was developed rigorously\\
\flagStyle{$\sf S_{\mbox{\scriptsize tested}}$}&Evidence that source code has been run with a clean build and tested\\
\flagStyle{$\sf S_{{\mbox{\scriptsize tools}}}$}&Evidence of any tool-based development\\
\flagStyle{$\sf S_{{\mbox{\scriptsize open source}}}$}&Team or open source development\\
\flagStyle{$\sf S_{{\mbox{\scriptsize otherSE}}}$}&Other evidence of good practice; see details in summary table\\
\flagStyle{$\sf C_0$}&Code has no non-trivial comments\\
\flagStyle{$\sf C_1$}&Code only has trivial comments (e.g., copyright)\\
\flagStyle{$\sf C_2$}&Helpful comments explaining code intent, rather than rephrasing the code\\
\flagStyle{$\sf C_{+}$}&Code has substantial, useful comments and documentation\\
\end{tabular}