### 2023-02-20 generated/info-for-main.tex
Sorting out .aux files not being deleted on make tidyup

\newlabel{supplementary-material-makefiles}{{11}{34}{DUMMYA}{DUMMYB}{DUMMYC}}
\newlabel{on-code-data-publication}{{12}{37}{DUMMYA}{DUMMYB}{DUMMYC}}
\newlabel{RAP-diagrams}{{12}{40}{DUMMYA}{DUMMYB}{DUMMYC}}
\newlabel{excelFigure}{{2}{40}{AAA}{BBB}{CCC}}
\newlabel{supplementary-Spiegelhalter-section}{{13}{40}{DUMMYA}{DUMMYB}{DUMMYC}}
\newlabel{survey-section}{{14}{43}{DUMMYA}{DUMMYB}{DUMMYC}}
\newlabel{supplementaryTableRepoSummary}{{4}{45}{DUMMYA}{DUMMYB}{DUMMYC}}
\newlabel{supplementary-journal-policies-section}{{14}{48}{DUMMYA}{DUMMYB}{DUMMYC}}
\newlabel{restrictive-policies}{{14}{52}{DUMMYA}{DUMMYB}{DUMMYC}}
\def\MaxMainPaperCitationNumber{0}