### 2023-02-20 generated/over-fitting-data.tex

\begin{center}\sf\begin{tabular}{|c|r|r|r|r|r|}\hline
$x={}$&1.00&1.10&2.20&2.60&4.50\\
$y={}$&1.50&2.70&4.90&5.70&8.20\\
\hline\end{tabular}
\end{center}

Showing both polynomials with coefficients to 2 decimal places, the linear least squares model to fit this data is:

$$
\hat{y}=0.49+1.80x
$$

and the Lagrange polynomial model (an exact fit to the 5 data points) is

$$
\hat{y}=-41.79+85.46x-56.30x^2+15.65x^3-1.51x^4
$$

The name of the relevant \emph{Mathematica\/} notebook file where all data and code for this section (and for generating the paper's section \ref{over-fit} and its figure \ref{fig-overfit}) can be found in \texttt{programs/over-fitting-section.nb}, which is included in the Git repository for the paper.