### 2023-05-12 generated/info-for-main.tex

\newlabel{supplementary-material-makefiles}{{9}{42}{DUMMYA}{DUMMYB}{DUMMYC}}
\newlabel{on-code-data-publication}{{10}{45}{DUMMYA}{DUMMYB}{DUMMYC}}
\newlabel{RAP-diagrams}{{10}{48}{DUMMYA}{DUMMYB}{DUMMYC}}
\newlabel{excelFigure}{{2}{48}{AAA}{BBB}{CCC}}
\newlabel{supplementary-Spiegelhalter-section}{{11}{48}{DUMMYA}{DUMMYB}{DUMMYC}}
\newlabel{survey-section}{{12}{51}{DUMMYA}{DUMMYB}{DUMMYC}}
\newlabel{supplementaryTableRepoSummary}{{4}{53}{DUMMYA}{DUMMYB}{DUMMYC}}
\newlabel{supplementary-journal-policies-section}{{12}{56}{DUMMYA}{DUMMYB}{DUMMYC}}
\newlabel{restrictive-policies}{{12}{60}{DUMMYA}{DUMMYB}{DUMMYC}}
\def\MaxMainPaperCitationNumber{126}