### 2023-05-12 generated/legend.tex

\begin{tabular}{lrp{4.5in}}
&\bfseries{Paper}&\\\bfseries{Flag}&\bfseries{count}&\bfseries{Meaning}\\ \hline 
\flagStyle{$\sf P_c$}&26&Journal has a code policy (see section~\ref{supplementary-journal-policies-section})\\ 
\flagStyle{$\sf P_{{\mbox{\scriptsize c-breach}}}$}&11&Paper breaches journal code policy (see section~\ref{supplementary-journal-policies-section})\\ 
\flagStyle{$\sf R_c$}&10&Paper uses a code repository (e.g., GitHub)\\ 
\flagStyle{$\sf R_{\mbox{\scriptsize c-empty}}$}&1&Code repository contains no code\\ 
\flagStyle{$\sf R_d$}&9&Paper uses a data repository (e.g., Dryad, Figshare, GitHub)\\ 
\flagStyle{$\sf S_{\mbox{\scriptsize NONE}}$}&12&No code available at all (note: code is not expected for standard models, systems or statistical methods)\\ 
\flagStyle{$\sf S_p$}&8&Paper says source code is available in principle\\ 
\flagStyle{$\sf S_{+}$}&12&Paper or URL provides source code\\ 
\flagStyle{$\sf S_{{\mbox{\scriptsize rigorous}}}$}&0&Evidence that source code was developed rigorously\\ 
\flagStyle{$\sf S_{\mbox{\scriptsize tested}}$}&0&Evidence that source code has been run with a clean build and tested\\ 
\flagStyle{$\sf S_{{\mbox{\scriptsize tools}}}$}&0&Evidence of any tool-based development\\ 
\flagStyle{$\sf S_{{\mbox{\scriptsize open source}}}$}&0&Team or open source development\\ 
\flagStyle{$\sf S_{{\mbox{\scriptsize otherSE}}}$}&1&Other evidence of good practice; see details in summary table\\ 
\flagStyle{$\sf C_0$}&10&Code only has comments unrelated to code intent (e.g., copyright)\\ 
\flagStyle{$\sf C_1$}&1&Code only has trivial or obvious comments\\ 
\flagStyle{$\sf C_2$}&3&Helpful comments explaining code intent, rather than rephrasing the code\\ 
\flagStyle{$\sf C_3$}&2&Code has substantial, useful comments, and documentation\\ 
\flagStyle{$\sf C_a$}&1&Code in repository uses assertions\\ 
\flagStyle{$\sf C_c$}&0&Code in repository uses pre- or post-conditions or similar\\ 
\end{tabular}