### 2023-05-12 generated/repos.tex

\begin{tabular}{|l|r|rrr@{\hskip .75ex}l|}\hline
&\multicolumn{1}{c|}{\bf PDF paper}&\multicolumn{4}{c|}{\bf Repository code \& data}\\ \cline{2-6}
\multicolumn{1}{|c|}{\bf Github repository}&\multicolumn{1}{c|}{\bf Number}&\multicolumn{1}{c}{\bf Number}&\multicolumn{1}{c}{\bf Code}&\multicolumn{2}{c|}{\bf Data}\\
\multicolumn{1}{|c|}{\bf and paper citation}&\multicolumn{1}{c|}{\bf of pages}&\multicolumn{1}{c}{\bf of files}&\multicolumn{1}{c}{\bf kLOC}&\multicolumn{2}{c|}{\bf bytes} \\ \hline\hline
\reponame{AI-CDSS-Cardiovascular-Silo}{AI-CDSS-Cardiovascular-Silo} & \csname pagelength-AI-CDSS-Cardiovascular-Silo\endcsname & \digitBlock{206} & 143 & 64&Mb \\
\reponame{blast-ct}{blast-ct} & \csname pagelength-blast-ct\endcsname & \digitBlock{54} & 3 & 238&Mb \\
\reponame{covid-sim}{covid-sim} & \csname pagelength-covid-sim\endcsname & \digitBlock{229} & 25 & 734&Mb \\
\reponame{lactModel}{lactModel} & \csname pagelength-lactModel\endcsname & \digitBlock{20} & 2 & 165&kb \\
\reponame{LRM}{LRM} & \csname pagelength-LRM\endcsname & \digitBlock{125} & 8 & 2&Mb \\
\reponame{manifold-ga}{manifold-ga} & \csname pagelength-manifold-ga\endcsname & \digitBlock{11} & 1 & \multicolumn{2}{c|}{---} \\
\reponame{MetricSelectionFramework}{MetricSelectionFramework} & \csname pagelength-MetricSelectionFramework\endcsname & \digitBlock{44} & 4 & 236&kb \\
\reponame{PENet}{PENet} & \csname pagelength-PENet\endcsname & \digitBlock{117} & 8 & 4&Mb \\
\reponame{philter-ucsf}{philter-ucsf} & \csname pagelength-philter-ucsf\endcsname & \digitBlock{1987} & 13 & 32&Mb \\
\reponame{PostoperativeOutcomes\_RiskNet}{PostoperativeOutcomes.RiskNet} & \csname pagelength-PostoperativeOutcomes.RiskNet\endcsname & \digitBlock{1} & --- & \multicolumn{2}{c|}{---} \\
\reponame{SiameseChange}{SiameseChange} & \csname pagelength-SiameseChange\endcsname & \digitBlock{5} & 1 & 1&kb \\
% Numbers below are rounded, but more precise averages are 254.454545 -- 18824.000000 -- 97771458.363636 = 1075486042 / 11
\hline \multicolumn{1}{|r|}{{\bf Average} ($N=11$)}&\makeAverage{\the\gitPages}{11}&254& 19& 98&Mb\\
\hline \end{tabular}