### 2023-05-12 generated/saved-files.tex

allGitRepos.sh & Shell script generated to download all available 10 GitHub repositories\\
assessments.tex & Main summary table for Supplementary Material\\
clone-date.tex & Running the download script also saves the download dates of GitHub clones\\
constants.tex & Generated \LaTeX\  definitions and constants\\
data-check.html & Convenient list of data sources with DOIs in HTML (enables checking of data just by clicking on HTML links)\\
data.csv & All data from \texttt{data.js} expressed in CSV for convenience\\
example-data.tex & \LaTeX\  example of JSON data\\
flagData.nb & Flags expressed as Mathematica code (facilitates ML or other analysis)\\
legend.tex & Legend for the main assessment table\\
open-doi.sh & Shell script to conveniently open all DOIs\\
page-lengths.tex & Page lengths and totals for Git data, and total pages of papers\\
saved-files.tex & This table of files generated by \texttt{data.js}\\
summary-table.tex & Short summary table\\
supplementary-references.tex & Reference list for Supplementary Material\\