### 2023-06-08 generated/make-help.tex
Fixing Elsevier quirks

% Edit makefile to change the selected make command choices, as follows:
% 1 - make all
% 2 - make archive
% 3 - make check
% 4 - make check-git
% 5 - make check-same
% 6 - make check-update
% 7 - make check-versions
% 8 - make data
% 9 - make expand
% 10 - make help
% 11 - make help-brief
% 12 - make mathematica
% 13 - make mathematica-open
% 14 - make one-file
% 15 - make pdf
% 16 - make push
% 17 - make readme
% 18 - make really-tidyup
% 19 - make tidyup
% 20 - make zip
% 21 - make zip-data

{\sf\begin{tabular}{rp{4.5in}}

\texttt{make all}&
      Download and analyze the data, then typeset the main files (\texttt{paper.pdf}, \texttt{appendix.pdf} and \texttt{all.pdf}) that import all the data, as well as make the self-contained expanded \texttt{.tex} (and \texttt{.PDF}) files that do not depend on the separate data files in \texttt{data/*}.\\
   \multicolumn{1}{l@{\vdots}}{}&\\

\texttt{make check-versions}&
      Check that you have the right software and software versions to run everything.\\
   \multicolumn{1}{l@{\vdots}}{}&\\

\texttt{make help}&
      Explain how to use \texttt{make}, and list all available options for using it.\\
\end{tabular}}